\documentclass[11pt]{article}

\newcommand{\school}{University of Michigan}
\newcommand{\schoolabbr}{U-M}
\newcommand{\dept}{U-M LSA Physics}

\usepackage[margin=1.5in]{geometry}

\title{\vspace{-5em}Academic Statement of Purpose\\
	\large \dept{}}

\author{Ethan Snyder}
\date{}

\begin{document}
\maketitle
\subsubsection*{My Path to the \school{}}
I'm a first generation college student raised in a family of six that always hovered around the poverty line. I watched my three older brothers barely scrape by in school, and so I was never encouraged to go above and beyond --- my parents were only concerned with me passing. Because of this, I owe much of my academic success to the opportunities and education I received in school. I'm grateful to have attended a well-funded school district with teachers who saw potential in me before I had seen it in myself, particularly in my early years. Without teachers and an education system that cared about me, I would have fallen behind before I found my passion for science. In elementary school, I was a student with behavioral issues. In middle school, I was put in accelerated classes, which enabled me to finish high school with a handful of college credits. My home was not an encouraging environment, so I'm incredibly lucky to have had a school district and teachers that enabled me to pursue higher education and get to where I am today.

During my freshman year at WMU, my parents divorced, and the divorce turned messy. In addition to being a full-time student and working part-time, I had to make time to appear in court as a witness, help my family clean and move out of our childhood home, and handle other family obligations. As my family and I distanced ourselves from my dad, we also parted with the last of our financial stability that he provided. This forced me into adulthood all at once and made it hard to prioritize my education. The effect of this period on my transcript is apparent, but also apparent is my resilience. When I emerged from this difficult period, I hit the ground running: I added a math major and a computer science minor to my physics and astronomy focus, moved into an apartment near campus, began assisting with research (both the accelerator work and the semiconductor research, the latter of which I continue today), and co-founded Physics Club.

These experiences have driven my decision to pursue a graduate degree at the University of Michigan, because I am drawn to Michigan's commitment to academic rigor, fostering diverse, underrepresented perspectives, and giving opportunity to those who need it most. These are values I've embraced throughout my journey, and values that have previously enabled me to get to where I am today. I look forward to building on my research foundation and becoming part of a community that prioritizes growth, resilience, and inclusivity.
\end{document}
