\documentclass[11pt]{article}

\usepackage{todonotes}
\newcommand{\school}{Michigan State University}
\newcommand{\schoolabbr}{MSU}
\newcommand{\dept}{Department of Physics and Astronomy}

\title{Academic Statement\\
	\large \dept{} at \school{}}

\author{Ethan Snyder}

\begin{document}
\maketitle
\subsubsection*{My Academic Experience}
All my life I had been certain that I wanted to study the stars. However, after a few years at university and over a year of assisting with semiconductor defect research, I grew to be fascinated by the physics of the small and the physics governing the technology that powers the modern world. This research affirmed my deep appreciation for computers, and since I've been certain that I want to work with simulations, models, and research fields applicable to computers. This is why I'm motivated to pursue condensed matter physics and quantum information science in graduate school. Graduate school is my next step, as it will allow me to further develop these interests, setting me up for my goal of researching semiconductors or quantum computing in the industry or as a professor. I see no better place to achieve this goal than \school{}, given the diversity and depth of research in these fields.

In my four years at Western Michigan University, I've been working towards Physics and Mathematics majors with Computer Science and Astronomy minors. This wasn't always the goal, as I realized halfway through my undergraduate experience that I enjoyed learning about and working with computers too much to not incorporate that into my education. I also realized that I could benefit from a stronger background in mathematics, more than just the calculus that is required in a physics degree. Despite the hefty courseload, I've been a dean's list student, been award a research scholarship through WMU's Undergraduate Research, Ingenuity and Creativity Summer Scholarship program (URICS), and most notably, in the spring of 2024, I received the Paul Rood Physics Scholarship from the Physics Department acknowledging exceptional performance within the department. I have certainly laid the foundations to be successful in computational research through both coursework and research experience, and I believe these accolades prove this.

By far my most extensive research experience at Western Michigan University is my research work with semiconductors with Dr.\@ O'Hara of the Physics Department, which, at the time of writing, I currently run on MSU's HPCC through ICER. The goal of this research is to calculate formation energies and model structural distortions in Ga and Al rich AlGaAs with a tellurium defect to corroborate and explain experimental findings. This research has familiarized me with solid state physics, semiconductors, structural defects, density functional theory, VASP, and different functionals like PBE, HSE, and VCA. I've also learned countless things about how to interface with computer clusters, how to move files around, and how to analyze data. I've attached a digital poster I presented at WMU's Lee Honors College as part of the previously mentioned URICS scholarship that summarizes much of my prep work and the work I'm currently doing.

In the spring of 2025, I will be beginning a semester project with Dr.\@ Korista of the Physics Department modeling stellar evolution in different classes of stars. This will provide me with a diversity of experiences with computational physics and data analysis across fields.
\todo{Add more to this, specifically about skills used and skills learned!}

During my time at Western, I've also become familiar with the Tandem Van de Graaff Accelerator in Rood Hall. While I have not operated it or officially participated in the research team that utilizes it, I have become familiar with how it is operated, what it is used for, and the data collection process, and have observed that research team often. As of this semester, I have also been spending time with the department's machine shop specialist learning to manufacture simple parts, learning machine shop etiquette, and how to communicate jobs effectively with machinists. Although my skills are best suited for theoretical, computational work, I believe my familiarity with the lab setting and part manufacturing has made me a well-rounded physicist and better able to help in experimental settings when necessary.

\subsubsection*{Why I choose \school{}}
\schoolabbr{} would provide me ample opportunity to build upon my knowledge and skills I've gained in computational condensed matter physics research, while also providing opportunities to explore a wide range of related topics. My confidence in navigating and running jobs on MSU's HPCC is a directly transferable skill that would enable me to quicker contribute to theoretical groups such as those of Dr.\@ Maghrebi and Dr.\@ Dykman's. This would give me the opportunity and experience needed to go forward into a career, onto the frontier of research.

The Maghrebi, Dykman, Becker, and Pollanen Group stand out to me the most. I'm most interested in the Maghrebi group, as I'll have over a year of continuous experience in condensed matter research by the time I graduate, and I'm deeply interested in continuing to learn about condensed matter physics. I will have a solid foundational understanding to contribute, and I'm excited about the possibility of contributing to this team while further developing my expertise in this field.

Leaning more into my interest of quantum information science and quantum computing, I can see myself working on Dr.\@ Dykman's research team. His work involving qubits and qubit frequency noise in particular is interesting, and I am confident that my thorough computational research experience will allow me to make meaningful contributions to this team. His work on electrons in helium and nanomechanics is also fascinating. Overall, Dr.\@ Dykman's research seems to be a steep learning curve, but I know the skills I've learned in undergrad have equipped me to take on such large tasks.

Dr.\@ Becker's group also really appeals to my interest in quantum information science and quantum computing, and I believe my extensive background in solid-state physics, atomic dopants and defects, would provide me a great starting point in understanding and contributing to this team. However, the extent of my experimental research experience is labs taken for credit and many hours around WMU's particle accelerator, although not explicitly operating it. I believe my familiarity with the lab setting and data analysis would enable me to learn from and contribute to the group's work. 

I'm eager to apply all that I've learned towards the fascinating research at \schoolabbr{}, particularly in areas that align with my interest in quantum computing and my experience with semiconductors. Thank you for your time and for considering my application. I look forward to the opportunity to learn from and contribute to your research teams.
\end{document}
